A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2020-07-01 07_Manuscript.Rmd
some small edits

The data for this study were obtained from through personal correspondence with colleagues and collaborating institutions.
The majority of the trials were undertaken by Professor Malcolm Riley and Dr Sue Thompson with the Queensland Department of Agriculture, Fisheries and Forestry (DAF). 
Trials undertaken in the 2013 season were conducted by the Northern Growers Alliance (NGA) [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013].
<!-- should check, Sue and Mal moved to USQ in 2015 I think. Therefore the trials probably started with USQ earlier --> The most recent trials starting from 2016 were undertaken by the University of Southern Queensland (USQ) with the assistance of Queensland DAF from 2016 to 2019. 
Trial data were in different formats with varying levels of information.
Some trial data reported means of each treatment, with all but one study reporting variance along with the mean. 
Raw data were obtained for 26 trials in total.

Details of trial data are presented within a research compendium which was prepared to supplement this publication [<!-- DOI to reference to compendium -->].

The criteria for inclusion of trial data in the meta-analysis required, a field trial testing fungicide efficacy for powdery mildew control on mungbean in the Grains Research and Development (GRDC) northern grains regions of eastern Australia, which grows the majority of Australia's summer crops.
Trial data had to include: the date when powdery mildew was first observed; disease incidence at the end of the growing season; fungicide application dates; and crop yield.
A total of 26 trials were collated and from these trials we selected the most frequently tested fungicides that had the same mode of action. 
The data were subset to only include treatments using demethylase inhibitor (DMI) fungicides, tebuconazole and propiconazole, to reduce confounding effects of fungicides with different modes of action. 
Subsequently the 26 trials were reduced to 17 using the same fungicide doses (Table 1).
With only a small number of trials and to increase the replication, we made no distinction between the tebuconazole and propiconazole in the analysis.

All trials included in this analysis were a randomised complete block designs and not previously published in peer-reviewed literature.
Some trials and results were reported online via the GRDC. 
While the main aim of all the trials collected for this study were to assess the efficacy of fungicide for control of powdery mildew, some trials included additional variables in the experimental design. 
These additional variables included: row spacing, fungicide active ingredient, fungicide dose, planting date, number of fungicide applications, cultivar and supplementary irrigation, in addition to trial location and season.
To account for this variation between trials, and the treatments within trials, a single 'trial' factor was created to describe the main variables; trial location, trial season, host genotype and row spacing.
Most of the studies which were used in this analysis assessed powdery mildew incidence on a 1 -- 9 ordinal scale (Table 2). 
However, the 2013 NGA trials (Table 1) reported incidence as the percentage of leaves covered by powdery mildew and the location of the infected leaves in the canopy; lower, middle or upper canopy.
Given the ordinal scale ranked plants on the percent of foliage up the plant showing the disease, it was strait-forward to calculate the equivalent incidence on the ordinal 1 -- 9 scale.

The timing of fungicide applications, with reference to when the disease first occurred in the trial, and number of applications were used to define a treatment variable. 
An 'Early' fungicide application, refers to a fungicide spray schedule that commences with the first spray prior to the disease being observed in the trial.
'Recommended' refers to a fungicide spray schedule which commences within three days following the first sign of disease, and 'late' treatments which commence four days or later than first sign.
The number of fungicide applications were binned in to two categories.
A spray schedule with a single spray, and a spray schedule with one or more sprays to a maximum of three applications.
The division of the treatments into these six categories created two categories with insufficient numbers of treatments to provide a reliable sample.
Single late sprays contained two replicates and multiple early sprays contained five replicates and were therefore removed from the subsequent meta-analysis.

@@

Mean grain yields and sample variance for each treatment was calculated for trials for which we had the raw data.
Sample variance for the 2012 Kingaroy study was calculated from the reported least squares statistic using a T-critical value of 1.697 and the approach reported in Ngugi et al. [-@Ngugi2011].
The T-critical value was determined from the 30 degrees of freedom using a t-distribution table.
For the remainder of the trial data, the treatment means and variance were reported, except for Gatton in 2012 and Emerald in 2015.
The Gatton and Emerald trials reported treatment means without accompanying variance and were thus removed prior to the analysis.

>>>>

The data for this study were obtained from through personal correspondence with colleagues and collaborating institutions.
The majority of the trials were undertaken by Professor Malcolm Riley and Dr Sue Thompson with the Queensland Department of Agriculture, Fisheries and Forestry (DAF). 
Trials undertaken in the 2013 season were conducted by the Northern Growers Alliance (NGA) [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013].
<!-- should check, Sue and Mal moved to USQ in 2015 I think. Therefore the trials probably started with USQ earlier --> The most recent trials starting from 2016 were undertaken by the University of Southern Queensland (USQ) with the assistance of Queensland DAF from 2016 to 2019. 
Trial data were in different formats with varying levels of information.
Some trial data reported means of each treatment, with all but one study reporting some form of variance along with the mean.

Twenty six trials in total were collated.

Details of trial data are presented within a research compendium which was prepared to supplement this publication [<!-- DOI to reference to compendium -->].
The criteria for inclusion of trial data in the meta-analysis required, a field trial testing fungicide efficacy for powdery mildew control on mungbean in the Grains Research and Development (GRDC) northern grains regions of eastern Australia, which grows the majority of Australia's summer crops.
Trial data had to include: the date when powdery mildew was first observed; disease incidence at the end of the growing season; fungicide application dates; the fungicide active ingredients; fungicide dose; and crop yield. 
The data were subset to only include treatments with the same mode of action; demethylase inhibitor (DMI) fungicides, tebuconazole and propiconazole were thus retained in the dataset. 
Subsequently the 26 trials were reduced to 16 using the same fungicide doses (Table 1). 
With only a small number of trials, and to increase the replication, we made no distinction between the tebuconazole and propiconazole in the analysis.

All trials included in this analysis were randomised complete block designs and not previously published in peer-reviewed literature.

While the main aim of all the trials collected for this study were to assess the efficacy of fungicide for control of powdery mildew, some trials tested additional variables in the experimental design.

These additional variables included: row spacing, fungicide active ingredient, fungicide dose, planting date, number of fungicide applications, cultivar and supplementary irrigation.
To account for the variation between trials, and the treatments within trials, a single 'trial' factor was created to describe the main variables; a trial reference code, trial location, trial season, host genotype, row spacing, and dose. 
Most of the studies which were used in this analysis assessed powdery mildew incidence on a 1 -- 9 ordinal scale (Table 2). 
However, the 2013 NGA trials (Table 1) reported incidence as the percentage of leaves covered by powdery mildew and the location of the infected leaves in the canopy; lower, middle or upper canopy.

Given the ordinal scale ranked plants on the percent of foliage up the plant showing the disease, a conversion to mean plot incidence on the ordinal 1 -- 9 scale was straight-forward.

The timing of fungicide application treatments, was defined in relation to the recorded date when powdery mildew first occurred in the trial. 
An 'Early' fungicide application treatment, referred to a fungicide spray schedule that commenced the first spray prior to the disease being observed in the trial. 
'Recommended' refers to a fungicide spray schedule which commences following the first sign of disease, within three days. 
'Late' treatments commence between 7 - 13 days after the first sign of disease. 
The number of fungicide applications were binned in to two categories.
A spray schedule with a single spray, and a spray schedule with two or three spray applications (Plus). 
The division of the treatments categorised: 13 Early treatments; 5 'EarlyPlus', 32 'Recommended'; 46 'RecommendedPlus'; 17 'Late'; and 20 'LatePlus.
'EarlyPlus' treatments were removed from the subsequent meta-analysis, due to an insufficient sample size.

@@

A categorical disease pressure factor was created by binning trials based on the mean AUDPC of the no spray control plot into two levels: 'low disease pressure' or 'high disease pressure'.
The two levels were separated via the median AUDPC (153.63) of the no spray control plots.

Mean grain yields for each treatment were either obtained directly from trial reports, or calculated from the raw data when available.
The sample variance was calculated from the raw data when available or the least squares statistic in trial reports. 
Sample variance for the 2012 Kingaroy study was calculated from the reported least squares statistic using a T-critical value of 1.697 and the approach reported in Ngugi et al. [-@Ngugi2011].
The T-critical value was determined from the 30 degrees of freedom using a t-distribution table. 
The 2012 Gatton and 2015 Emerald trials reported treatment means without accompanying variance and were thus removed from the dataset  prior to the analysis.

This analysis use the grain yield standardised mean difference (SMD) as the response variable. 
SMD were chosen to reduce the inherent variability between trials due to location and seasonal effects on grain yields.  
$SMD = \frac{Y_i-\overline{Y_c}}{\sigma_p}$ [@HigginsGreen]  
Where $Y_i$ is the treatment mean, $Y_c$ is the no spray control treatment mean and $\sigma_p$ is the trial pooled standard deviation.

Similar to Paul et. al. [-@Paul2007], we fitted a multivariate random effects model  
$Y_i \sim N(\mu,s_i^2 + \sigma^2)$  
where $Y_i$ is the response, a vector of the SMD grain yield for each treatment in the $i$th trial over total number ($K$) of trials($i$  = 1, ...$K$). 
$Y_i$ is estimated from a normal distribution with the mean $\mu$, sample variance of $Si_^2$ and a between trial variance of $\sigma^2$.  
**Moderator variables.** The random effects meta-analysis was expanded to evaluate if the residual heterogeneity could be explained by spray schedule ($SPRAY_{Sched}$). 
The intercept was omitted from the model so each treatment would be reported against zero, which represents the control mean. 
Linear contrasts were used to compare the mean effect sizes between each spray schedule treatment and their respecting standard errors and confidence intervals.
